---
title: "Total Covid Cases in the World"
author: "Kenneth Fajardo"
date: "6/5/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This presentation demonstrates the use of the Leaflet package in R. The data that we will be using here is the COVID data made available by [Our World In Data]("https://github.com/owid/covid-19-data/tree/master/public/data/"), focusing on COVID activity in Asia, as of 2021-06-03. Geodata are pulled from \@johan's [repo]("https://github.com/johan/world.geo.json").

## Initialization
Load the needed libraries
```{r lib, message=FALSE}
library(leaflet) # for generating interactive maps
library(geojsonio) # for retrieving geodata
library(dplyr) # for cleaning and manipulating data
library(RColorBrewer) # for coloring the choropleth
```

Download the data.
```{r dl, eval=FALSE}
download.file("https://raw.githubusercontent.com/owid/covid-19-data/
              master/public/data/owid-covid-data.csv", "covid.csv")
```

***
Read the data.
```{r read}
covid <- read.csv("covid.csv")
WorldCountry <-geojsonio::geojson_read("./data/countries.geo.json",
                                       what = "sp")
```

Select columns describing total cases in Asia per country.
```{r total, message=FALSE}
covid <- covid[,grep("total|iso|location|date", names(covid))]
covid <- covid %>% filter(date == "2021-06-03") %>% 
            group_by(iso_code, location) %>% 
            summarise(total = round(sum(total_cases ,na.rm=TRUE), 0))
covid <- covid[-grep("OWID", covid$iso_code),]
```

## Generate choropleth
**Total number of COVID cases per country as of 2021-06-03**
```{r geodata, echo=FALSE, message=FALSE, comment=""}
# Create base map
data_Map <- WorldCountry[WorldCountry$id %in% covid$iso_code, ]
covid <- covid[covid$iso_code %in% data_Map@data$id,]
Map <- leaflet(data_Map) %>% addTiles() %>% addPolygons()

# Set bin and color for choropleth map
bins <- c(0, 1000, 10000, 100000, 500000, 1000000, 5000000, 10000000, 20000000, Inf)
pal <- colorBin(rev(brewer.pal(9, "Reds")), domain = covid$total, bins = bins, reverse = TRUE)

# Set labels
labels <- sprintf(
  "<strong>%s</strong> <br/>%s Total Cases <sup></sup>",
  data_Map@data$name, format(covid$total, scientific = FALSE, big.mark = ",")) %>%
    lapply(htmltools::HTML)

# Customize map
Map %>% addPolygons(
  fillColor = ~pal(covid$total),
  weight = 1,
  stroke = FALSE,
  color = "white",
  opacity = 0.2,
  dashArray = '2',
  fillOpacity = 0.3,
  highlight = highlightOptions(
    weight = 2,
    color = "red",
    dashArray = "",
    fillOpacity = 0.4,
    bringToFront = TRUE),
  label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto")
) %>% addLegend(pal = pal, values = covid$total, opacity = 0.7, title = "Covid Cases",
  position = "bottomright")
```

<font size="2">Note that there are **issues** in mapping COVID cases for the **US**. However, the goal of this project is to experiment on Leaflet.</font>